Fast Hyperparameter Selection for Graph Kernels via Subsampling and\\Multiple Kernel Learning

### 2016-11-26 paper.tex
F

Model selection is one of the most computationally expensive tasks in a
machine learning application. When dealing with kernel methods for structures,
the choice with the largest impact on the overall performance is the selection
of the feature bias, i.e. the choice of the concrete kernel for structures.
Each kernel in turn exposes several hyper-parameters which also need to be
fine tuned.  Multiple Kernel Learning offers a way to approach this
computational bottleneck by generating a combination of different kernels
under different parametric settings. However, this solution still requires the
computation of many large kernel matrices. In this paper we propose a method
to efficiently select just a small number of kernels without loss in
predictive performance.

@@

Many application domains can be naturally encoded in a structured form. In
chemo- and bioinformatics for example, molecules and polymers can be encoded
as graphs with atoms as nodes and chemical bonds as edges. After a
segmentation pre-processing step, images can also be represented as networks
of adjacent objects or regions. Working on these domains employing traditional
feature engineering is possible, but it is generally time-consuming (every
different application requires a novel feature design by some expert) and
information loss is often unavoidable. Direct processing of structures is
possible using a kernelized approach, where a kernel function can be devised
for each generic data type (for example sequences, trees or graphs). This in
turn allows to use standard learning algorithms (e.g. SVM) to solve the
predictive task.

Model selection is one of the most computationally expensive tasks in a
machine learning application. When dealing with kernel methods for structures,
the choice with the largest impact on the overall performance is the selection
of the feature bias, i.e. the choice of the concrete kernel for structures.
Each kernel in turn exposes several hyper-parameters which also need to be
fine tuned. Finally, the downstream learning algorithm has its own hyper-
parameters. A common approach to hyper-parameters optimization is to fix
a-priori  a \textit{parameters grid}, i.e. a set of possible values for each
parameter. Then, the predictive performance of the learning algorithm is
evaluated for each possible parameters combination ({\em grid-search}
strategy), which is possibly a very time consuming step.  A more efficient
approach has been proposed in~\cite{FabioAiolliMicheleDonini2015,Massimo2016},
where instead of selecting the best parameters configuration, Multiple Kernel
Learning (MKL) was used for combining different kernel matrices in a single
learning procedure. While more efficient, this approach does not scale to very
large datset since a kernel matrix has to be computed for each kernel and for
each of its hyper-parameters. In this paper we propose a new approach that,
based on sub-sampling and MKL, is able to speedup the parameter selection
phase by orders of magnitude on real-world datasets  without loss in
predictive performance.

@@

\section{Discussion and conclusion}
In the context of MKL, we showed a procedure to learn the kernels combination vector using a subset of training data, and use this vector to chose the top kernels for a given problem. Then we propose three methods to combine top kernels quickly, preserving an high AUC score.
As showed in the results table, the percentage of subsampling and top kernels selected (10\%) is a good compromise between time complexity and high AUC score. Furthermore, due to the low number of selected top kernels, we can assume that all of them are good kernels for our problem, so we can combine them effectively using the simple summation, avoiding to use the EasyMKL algorithms in the second learning step.\\
Finally, our method is able to reduce effectively the time complexity, making the MKL approach usable for huge datasets.

>>>>

Model selection is one of the most computationally expensive tasks in a
machine learning application. When dealing with kernel methods for structures,
the choice with the largest impact on the overall performance is the selection
of the feature bias, i.e. the choice of the concrete kernel for structures.
Each kernel in turn exposes several hyper-parameters which also need to be
fine tuned.  Multiple Kernel Learning offers a way to approach this
computational bottleneck by generating a combination of different kernels
under different parametric settings. However, this solution still requires the
computation of many large kernel matrices. In this paper we propose a method
to efficiently select a small number of kernels on a subset of the original
data, gaining a dramatic reduction in the runtime without a significant loss
of predictive performance.

@@

Many application domains can be naturally encoded in a structured form. In
chemo- and bioinformatics for example, molecules and polymers can be encoded
as graphs with atoms as nodes and chemical bonds as edges. After a
segmentation pre-processing step, images can also be represented as networks
of adjacent objects or regions. Working on these domains employing traditional
feature engineering is possible, but it is generally time-consuming (every
different application requires a novel feature design by some expert) and
information loss is often unavoidable. Instead, a direct processing of
structures can be obtained using kernelized approaches, where a kernel
function can be devised for each generic data type (for example sequences,
trees or graphs). This in turn allows to use standard learning algorithms
(e.g. SVM) to solve the desired predictive tasks.

Model selection is one of the most computationally expensive tasks in
machine learning applications. When dealing with kernel methods for structures,
the choice with the largest impact on the overall performance is the selection
of the feature bias, i.e. the choice of the concrete kernel for structures.
Each kernel in turn exposes several hyper-parameters which also need to be
fine tuned. Finally, the downstream learning algorithm has its own hyper-
parameters. A common approach to hyper-parameters optimization is to fix
a-priori  a \textit{parameters grid}, i.e. a set of possible values for each
parameter. Then, the predictive performance of the learning algorithm is
evaluated for each possible parameters combination ({\em grid-search}
strategy), which is possibly a very time consuming step.  A more efficient
approach has been proposed in~\cite{FabioAiolliMicheleDonini2015,Massimo2016},
where instead of selecting the best parameters configuration, Multiple Kernel
Learning (MKL) was used for combining different kernel matrices in a single
learning procedure. While more efficient, this approach does not scale to very
large datset since a kernel matrix has to be computed for each kernel and for
each of its hyper-parameters. In this paper we propose a new approach that,
based on sub-sampling and MKL, is able to speedup the parameter selection
phase by orders of magnitude on real-world datasets  without loss in
predictive performance.

@@

\section{Discussion and conclusion}

We have presented an efficient approach to model selection for graph kernels
based on the multiple kernel learning framework. We have shown how working on
small subsets (up to 10\%) of the data yields results that are stable and
accurate enough to select the optimal combination of kernels via EasyMKL.
We have investigated three methods to efficiently combine a subset of
kernels while preserving at the same time an high AUC score for the downstream
machine learning algorithms. Finally, we showed how we can use the simple
summation of the selected kernels, avoiding the expensive step of running MKL
on the full training set, to scale the MKL approach to very large
datasets.

In future work we will extend the library of graph kernels to obtain a
comprehensive and easy to use software package.
